---
title: "MA7419 Fundamental Of Data Science - Coursework 2, 2025/26 , Group 44"
output: html_document
date: "2025-11-24"
---

```{r setup, include=FALSE}

library(tidyverse)
library(scales)

theme_set(
  theme_minimal(base_size = 14) +
    theme(
      plot.title    = element_text(face = "bold", size = 18),
      plot.subtitle = element_text(size = 12),
      panel.grid.minor = element_blank(),
      legend.position  = "bottom"
    )
)


```

---

## Research Questions

- **RQ1:** How have Brazil's CO₂ emissions per capita changed over time, and how do they compare to the global average?  
  *Approach:* Trend plots + t-test.

- **RQ2:** How dependent is Brazil on different electricity sources (coal, solar, wind, hydro), and how are these related to its CO₂ emissions?  
  *Approach:* Correlation analysis + ANOVA + line charts.

- **RQ3:** Has Brazil’s CO₂ emission pattern changed significantly before and after the year 2000?  
  *Approach:* Before/after plot + chi-square test.

- **RQ4:** How have Brazil’s CO₂ emissions and renewables changed relative to India and the USA?  
  *Approach:* Comparison plots (Brazil, India, USA).
  
- **RQ5:** How do Brazil’s renewable energy investments compare to other BRICS nations?
  *Approach:* comparative multi-country statistical and visual analytics
---

# 1. Load datasets

```{r}
co2 <- read_csv(
("https://raw.githubusercontent.com/owid/co2-data/master/owid-co2-data.csv"),
show_col_types = FALSE
)

energy <- read_csv(
("https://raw.githubusercontent.com/owid/energy-data/master/owid-energy-data.csv"),
show_col_types = FALSE
)

colnames(co2)
colnames(energy)

```
```{r}
co2_clean <- co2 %>%
  rename(country = country, year = year) %>%
  filter(!is.na(year))

co2_use <- co2_clean %>%
  select(country, year, co2, co2_per_capita, land_use_change_co2)

energy_clean <- energy %>%
  rename(country = country, year = year) %>%
  filter(!is.na(year))

energy_use <- energy_clean %>%
  select(
    country, year,
    solar_electricity, wind_electricity, hydro_electricity,
    nuclear_electricity, coal_electricity, gas_electricity, oil_electricity
  )

merged_all <- inner_join(
  co2_use,
  energy_use,
  by = c("country", "year")
)

countries_focus <- c("Brazil", "India", "United States")
merged_focus <- merged_all %>% filter(country %in% countries_focus)

merged_brazil <- merged_all %>% filter(country == "Brazil")


global_data <- merged_all %>%
  group_by(year) %>%
  summarise(
    global_co2_per_capita = mean(co2_per_capita, na.rm = TRUE),
    global_renewables = mean(solar_electricity + wind_electricity + hydro_electricity, na.rm = TRUE)
  )

```
#RQ1 – Brazil’s CO₂ per capita over time vs global

```{r}
brazil_global <- merged_brazil %>%
  left_join(
    global_data %>%
      select(year, global_co2_per_capita),
    by = "year"
  ) %>%
  drop_na(co2_per_capita, global_co2_per_capita)

last_row <- brazil_global %>%
  filter(year == max(year, na.rm = TRUE))

ggplot(brazil_global, aes(x = year)) +
  geom_ribbon(aes(ymin = pmin(co2_per_capita, global_co2_per_capita),
                  ymax = pmax(co2_per_capita, global_co2_per_capita)),
              fill = "#1b4332", alpha = 0.4) +
  
  geom_line(aes(y = global_co2_per_capita, colour = "Global average"),
            linewidth = 1.1, linetype = "dashed", alpha = 0.9) +
  
  geom_line(aes(y = co2_per_capita, colour = "Brazil"),
            linewidth = 1.6) +
  
  geom_point(data = last_row,
             aes(y = co2_per_capita, colour = "Brazil"),
             size = 3) +
  geom_point(data = last_row,
             aes(y = global_co2_per_capita, colour = "Global average"),
             size = 3) +
  
  geom_text(
    data = last_row,
    aes(y = co2_per_capita,
        label = sprintf("Brazil: %.2f", co2_per_capita)),
    hjust = -0.1, vjust = -0.5, size = 3.5, colour = "#52ffb8"
  ) +
  geom_text(
    data = last_row,
    aes(y = global_co2_per_capita,
        label = sprintf("Global: %.2f", global_co2_per_capita)),
    hjust = -0.1, vjust = 1.2, size = 3.5, colour = "grey80"
  ) +
  
  scale_colour_manual(
    values = c(
      "Brazil"         = "#52ffb8",  # neon green
      "Global average" = "#ffdd57"   # warm yellow
    )
  ) +
  scale_x_continuous(breaks = scales::pretty_breaks(8)) +
  
  labs(
    title    = "Brazil vs Global CO₂ Emissions per Capita",
    subtitle = "Dark-mode view • Brazil's per-person emissions vs world average",
    x        = "Year",
    y        = "CO₂ per capita (tonnes)",
    colour   = ""
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.background  = element_rect(fill = "#081018", colour = NA),
    panel.background = element_rect(fill = "#081018", colour = NA),
    panel.grid.major = element_line(colour = "#1f2933"),
    panel.grid.minor = element_blank(),
    text             = element_text(colour = "grey90"),
    axis.text        = element_text(colour = "grey80"),
    axis.title       = element_text(colour = "grey90"),
    plot.title       = element_text(colour = "white", face = "bold", size = 18),
    plot.subtitle    = element_text(colour = "grey80"),
    legend.background= element_rect(fill = "#081018", colour = NA),
    legend.key       = element_rect(fill = "#081018", colour = NA),
    legend.position  = "top"
  ) +
  coord_cartesian(clip = "off")

```
#Decadal averages bar chart

```{r}

brazil_decade <- brazil_global %>%
  mutate(decade = floor(year / 10) * 10) %>%
  group_by(decade) %>%
  summarise(
    brazil_mean = mean(co2_per_capita,         na.rm = TRUE),
    global_mean = mean(global_co2_per_capita,  na.rm = TRUE),
    .groups = "drop"
  ) %>%
  pivot_longer(
    cols      = c(brazil_mean, global_mean),
    names_to  = "series",
    values_to = "value"
  ) %>%
  mutate(
    series = recode(series,
                    "brazil_mean" = "Brazil",
                    "global_mean" = "Global")
  )

ggplot(brazil_decade,
       aes(x = factor(decade), y = value, fill = series)) +
  
  geom_col(position = position_dodge(width = 0.7), width = 0.55,
           colour = "white", alpha = 0.9) +
  
  geom_text(aes(label = sprintf("%.2f", value)),
            position = position_dodge(width = 0.7),
            vjust = -0.6, size = 3.8, colour = "white") +
  
  
  
  labs(
    title    = "Decadal Average CO₂ Emissions per Capita",
    subtitle = "Brazil vs Global Comparison Across Decades",
    x        = "Decade",
    y        = "Average CO₂ per Capita (tonnes)",
    fill     = ""
  ) +

  expand_limits(y = max(brazil_decade$value) * 1.15) +

  theme_minimal(base_size = 14) +
  theme(
    plot.background  = element_rect(fill = "#081018", colour = NA),
    panel.background = element_rect(fill = "#081018", colour = NA),
    panel.grid.major = element_line(colour = "#1f2933"),
    panel.grid.minor = element_blank(),
    axis.text        = element_text(colour = "grey80"),
    axis.title       = element_text(colour = "white"),
    plot.title       = element_text(colour = "white", size = 18, face = "bold"),
    plot.subtitle    = element_text(colour = "grey80"),
    legend.position  = "top",
    legend.text      = element_text(colour = "white"),
    legend.background= element_rect(fill = "#081018", colour = NA)
  )



```

#RQ2 – Hydropower dependence vs other renewables
#a) Trend lines: Hydro vs Solar vs Wind
# Heatmap of renewables by decade – “energy intensity grid”

```{r}

brazil_renew <- energy %>%
  filter(country == "Brazil") %>%
  transmute(
    year,
    solar = solar_electricity,
    wind  = wind_electricity,
    hydro = hydro_electricity
  ) %>%
  pivot_longer(
    cols      = -year,
    names_to  = "source",
    values_to = "value"
  ) %>%
  filter(!is.na(value))


brazil_renew_decade <- brazil_renew %>%
  mutate(decade = floor(year / 10) * 10) %>%
  group_by(decade, source) %>%
  summarise(
    mean_value = mean(value, na.rm = TRUE),
    .groups    = "drop"
  )

ggplot(brazil_renew_decade, aes(x = factor(decade), y = source, fill = mean_value)) +
  geom_tile(color = "#2C2C2C", linewidth = 0.6) +
  
  scale_fill_gradient(
    low  = "#1B2631",   
    high = "#85C1E9",   
    name = "Avg TWh"
  ) +
  
  labs(
    title = "Brazil Renewable Electricity by Decade",
    subtitle = "Hydro dominates earlier decades; wind and solar rise after 2000",
    x = "Decade",
    y = "Energy Source"
  ) +
  
  # DARK THEME
  theme_minimal(base_size = 14) +
  theme(
    plot.background   = element_rect(fill = "#121212", color = NA),
    panel.background  = element_rect(fill = "#121212", color = NA),
    legend.background = element_rect(fill = "#121212", color = NA),
    legend.key        = element_rect(fill = "#121212", color = NA),
    
    axis.text         = element_text(color = "white"),
    axis.title        = element_text(color = "white"),
    
    plot.title        = element_text(color = "white", face = "bold", size = 18),
    plot.subtitle     = element_text(color = "grey80", size = 12),
    
    legend.title      = element_text(color = "white"),
    legend.text       = element_text(color = "white"),
    
    panel.grid        = element_blank()
  )

```



#b) ANOVA: mean output differences between sources

```{r}
anova_brazil <- aov(value ~ source, data = brazil_renew)
summary(anova_brazil)

```
#c) Correlation: Hydro vs Other Renewables (Solar + Wind)

```{r}
brazil_renew_corr <- merged_brazil %>%
  transmute(
    year,
    hydro        = hydro_electricity,
    other_renew  = solar_electricity + wind_electricity
  ) %>%
  drop_na(hydro, other_renew)

cor.test(brazil_renew_corr$hydro, brazil_renew_corr$other_renew)


```
#🌳 RQ3 – “Deforestation” trend via land-use change CO₂ (before vs after 2000)
#a) Trend graph (with 2000 marked)

```{r brazil_luc_data}
brazil_luc <- co2 %>%
  filter(country == "Brazil") %>%
  select(year, land_use_change_co2) %>%
  filter(!is.na(land_use_change_co2))

theme_dark_glow <- theme_minimal(base_size = 14) +
  theme(
    plot.background   = element_rect(fill = "#0D0D0D", color = NA),
    panel.background  = element_rect(fill = "#0D0D0D", color = NA),
    panel.grid.major  = element_line(color = "#2A2A2A", size = 0.3),
    panel.grid.minor  = element_blank(),

    axis.text         = element_text(color = "grey85", size = 13),
    axis.title        = element_text(color = "white", size = 14, face = "bold"),

    plot.title        = element_text(color = "white", face = "bold", size = 20),
    plot.subtitle     = element_text(color = "grey80", size = 13),
    plot.caption      = element_text(color = "grey60", size = 10),
  )

ggplot(brazil_luc, aes(x = year, y = land_use_change_co2)) +

  geom_line(color = "#FF3B3B40", linewidth = 5, alpha = 0.35) +

  geom_line(color = "#FF5252", linewidth = 1.6) +

  geom_smooth(color = "#FFA07A", linewidth = 1.2, se = FALSE, span = 0.3) +

  annotate("rect",
           xmin = 1995, xmax = 2010,
           ymin = -Inf, ymax = Inf,
           alpha = 0.05, fill = "#FF0000") +
  annotate("text",
           x = 2002, y = max(brazil_luc$land_use_change_co2, na.rm = TRUE),
           label = "High Deforestation Period", 
           color = "red", size = 4, fontface = "bold") +

  labs(
    title    = "Brazil Land-Use Change CO₂ Emissions",
    subtitle = "Deforestation and land-use change emissions highlighted with trend smoothing",
    x        = "Year",
    y        = "Land-use CO₂ (million tonnes)",
    caption  = "Source: OWID CO₂ Dataset"
  ) +
  theme_dark_glow

```
#b) Chi-square: Before vs After 2000 (High vs Low deforestation emissions)
```{r}

brazil_luc_period <- brazil_luc %>%
  mutate(
    period   = ifelse(year < 2000, "Before 2000", "After 2000"),
    category = ifelse(
      land_use_change_co2 > median(land_use_change_co2, na.rm = TRUE),
      "High", "Low"
    )
  )

luc_tab <- table(brazil_luc_period$period, brazil_luc_period$category)
luc_tab

chi_sq_luc <- chisq.test(luc_tab)
chi_sq_luc
```

#🌐 RQ4 – Brazil vs India vs USA: CO₂ & renewables
#a) CO₂ per capita comparison
```{r co2_compare_dark_facet}

compare_countries <- c("Brazil", "India", "United States")

co2_compare <- co2 %>%
  filter(country %in% compare_countries) %>%
  select(country, year, co2_per_capita) %>%
  filter(!is.na(co2_per_capita))

theme_dark_custom <- theme_minimal(base_size = 14) +
  theme(
    plot.background  = element_rect(fill = "#0d1117", colour = NA),
    panel.background = element_rect(fill = "#0d1117", colour = NA),
    panel.grid.major = element_line(colour = "#2d333b"),
    panel.grid.minor = element_blank(),
    axis.text        = element_text(colour = "grey85"),
    axis.title       = element_text(colour = "grey90"),
    strip.background = element_rect(fill = "#161b22", colour = NA),
    strip.text       = element_text(colour = "grey90", face = "bold"),
    plot.title       = element_text(colour = "white", face = "bold", size = 20),
    plot.subtitle    = element_text(colour = "grey80"),
    legend.background = element_rect(fill = "#0d1117"),
    legend.key       = element_rect(fill = "#0d1117"),
    legend.text      = element_text(colour = "grey85"),
    legend.title     = element_text(colour = "white")
  )

ggplot(co2_compare, aes(x = year, y = co2_per_capita, colour = country)) +
  geom_line(linewidth = 1.3) +
  facet_wrap(~ country, scales = "free_y") +
  scale_colour_manual(values = c(
    "Brazil"         = "#00e676",  
    "India"          = "#ffa726",  
    "United States"  = "#42a5f5"  
  )) +
  labs(
    title    = "Country-wise CO₂ Emissions per Capita",
    subtitle = "Brazil, India and the United States",
    x        = "Year",
    y        = "CO₂ per capita (tonnes)"
  ) +
  theme_dark_custom +
  theme(legend.position = "none")

```

#b) Renewable electricity comparison (solar + wind + hydro)
```{r renew_compare_boxplot, message=FALSE, warning=FALSE}

compare_countries <- c("Brazil", "India", "United States")

renew_compare <- energy %>%
  filter(country %in% compare_countries) %>%
  mutate(
    Entity     = country,
    Year       = year,
    renewables = solar_electricity + wind_electricity + hydro_electricity
  ) %>%
  select(Entity, Year, renewables) %>%
  filter(!is.na(renewables))

renew_compare <- renew_compare %>%
  mutate(Entity = reorder(Entity, renewables, FUN = median, decreasing = TRUE))

theme_dark_custom <- theme_minimal(base_size = 12) +
  theme(
    plot.background  = element_rect(fill = "#0d1117", colour = NA),
    panel.background = element_rect(fill = "#0d1117", colour = NA),
    panel.grid.major = element_line(colour = "#2d333b"),
    panel.grid.minor = element_blank(),
    axis.text        = element_text(colour = "grey85"),
    axis.title       = element_text(colour = "grey90"),
    plot.title       = element_text(colour = "white", face = "bold", size = 15),
    plot.subtitle    = element_text(colour = "grey80"),
    legend.background = element_rect(fill = "#0d1117"),
    legend.key       = element_rect(fill = "#0d1117"),
    legend.text      = element_text(colour = "grey85"),
    legend.title     = element_text(colour = "white")
  )

ggplot(renew_compare, aes(x = Entity, y = renewables, fill = Entity)) +
  geom_violin(alpha = 0.45, colour = NA, width = 0.7) +
  
  geom_boxplot(
    width = 0.25,
    alpha = 0.8,
    outlier.shape = NA,
    colour = "grey95",
    linewidth = 0.4
  ) +
  
  geom_jitter(
    width  = 0.15,
    alpha  = 0.9,
    colour = "white",
    size   = 1.5
  ) +
  
  scale_fill_manual(values = c(
    "Brazil"        = "#00e676",  # neon green
    "India"         = "#ffa726",  # warm orange
    "United States" = "#42a5f5"   # soft blue
  )) +
  
  scale_y_continuous(labels = comma) +
  
  labs(
    title    = "Renewable Electricity (TWh): Brazil vs India vs USA",
    subtitle = "Solar + wind + hydro – distribution across years (violin + boxplot + white points)",
    x        = "Country",
    y        = "Renewable Electricity (TWh)",
    fill     = "Country"
  ) +
  theme_dark_custom +
  theme(legend.position = "none")


```
```{r brics_renewables_analysis, message=FALSE, warning=FALSE}

#  Focus on BRICS countries --------------------------------------------
brics <- c("Brazil", "Russia", "India", "China", "South Africa")

brics_energy <- energy %>%
  filter(country %in% brics) %>%
  # Robust renewables TWh variable:
  mutate(
    renewables_twh = case_when(
      !is.na(renewables_electricity) ~ renewables_electricity,
      TRUE ~ coalesce(
        solar_electricity, 0
      ) + coalesce(
        wind_electricity, 0
      ) + coalesce(
        hydro_electricity, 0
      )
    )
  ) %>%
  filter(!is.na(year))

#  Summary table: latest year for each BRICS country -------------------
summary_latest <- brics_energy %>%
  # Keep only years where renewables_share_energy isn't NA
  filter(!is.na(renewables_share_energy)) %>%
  group_by(country) %>%
  slice_max(year, n = 1, with_ties = FALSE) %>%
  ungroup() %>%
  select(
    country,
    year,
    renewables_share_energy,
    renewables_twh,
    primary_energy_consumption
  ) %>%
  arrange(desc(renewables_share_energy))

cat("\n=== Latest Available Renewables Snapshot (BRICS) ===\n")
print(
  summary_latest %>%
    mutate(
      renewables_share_energy = round(renewables_share_energy, 1),
      renewables_twh          = round(renewables_twh, 1),
      primary_energy_consumption = round(primary_energy_consumption, 1)
    )
)

#  Decade-level summary for heatmap ------------------------------------
brics_decade <- brics_energy %>%
  mutate(decade = floor(year / 10) * 10) %>%
  group_by(country, decade) %>%
  summarise(
    renewables_share_mean = mean(renewables_share_energy, na.rm = TRUE),
    renewables_twh_mean   = mean(renewables_twh, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  filter(!is.na(renewables_share_mean))

cat("\n=== Decade-Average Renewables Share (BRICS) ===\n")
print(
  brics_decade %>%
    arrange(country, decade) %>%
    mutate(
      renewables_share_mean = round(renewables_share_mean, 1),
      renewables_twh_mean   = round(renewables_twh_mean, 1)
    )
)

#  Line plot: renewables share over time (BRICS) -----------------------

theme_brics_dark <- theme_minimal(base_size = 14) +
  theme(
    plot.background  = element_rect(fill = "#0d1117", colour = NA),
    panel.background = element_rect(fill = "#0d1117", colour = NA),
    panel.grid.major = element_line(colour = "#2d333b"),
    panel.grid.minor = element_blank(),
    axis.text        = element_text(colour = "grey85"),
    axis.title       = element_text(colour = "grey90"),
    strip.background = element_rect(fill = "#161b22", colour = NA),
    strip.text       = element_text(colour = "grey90", face = "bold"),
    plot.title       = element_text(colour = "white", face = "bold", size = 18),
    plot.subtitle    = element_text(colour = "grey80"),
    legend.background = element_rect(fill = "#0d1117"),
    legend.key       = element_rect(fill = "#0d1117"),
    legend.text      = element_text(colour = "grey85"),
    legend.title     = element_text(colour = "white")
  )

ggplot(
  brics_energy %>% filter(!is.na(renewables_share_energy)),
  aes(x = year, y = renewables_share_energy, colour = country)
) +
  geom_line(linewidth = 1.3, alpha = 0.9) +
  scale_colour_manual(values = c(
    "Brazil"        = "#00e676",  # neon green
    "Russia"        = "#ff7043",  # orange
    "India"         = "#ffde59",  # yellow
    "China"         = "#42a5f5",  # blue
    "South Africa"  = "#ab47bc"   # purple
  )) +
  labs(
    title    = "Renewables Share of Primary Energy – BRICS",
    subtitle = "Brazil stands out with a consistently higher renewable share",
    x        = "Year",
    y        = "Renewables share of energy (%)",
    colour   = "Country"
  ) +
  theme_brics_dark

#  Heatmap: renewables by decade & country -----------------------------

ggplot(
  brics_decade,
  aes(
    x    = factor(decade),
    y    = country,
    fill = renewables_share_mean
  )
) +
  geom_tile(colour = "#2d333b", linewidth = 0.5) +
  scale_fill_gradient(
    low  = "#13293D",
    high = "#3dfc8f",
    name = "Avg renewables\nshare (%)"
  ) +
  labs(
    title    = "BRICS Renewable Energy by Decade",
    subtitle = "Average renewables share of primary energy by decade",
    x        = "Decade",
    y        = "Country"
  ) +
  theme_brics_dark

# MAP  — BRICS Renewable Share (%) -----------------------------

library(sf)
library(rnaturalearth)
library(rnaturalearthdata)  


world <- ne_countries(scale = "medium", returnclass = "sf")

brics_map <- world %>%
  filter(name %in% c("Brazil", "Russia", "India", "China", "South Africa")) %>%
  left_join(summary_latest, by = c("name" = "country"))   # <-- closed )

ggplot(brics_map) +
  geom_sf(aes(fill = renewables_share_energy), color = "grey30", size = 0.3) +
  scale_fill_gradient(
    low  = "#13293D",
    high = "#3dfc8f",
    name = "Renewables Share (%)"
  ) +
  theme_minimal() +
  theme(
    plot.background  = element_rect(fill = "#0d1117", colour = NA),
    panel.background = element_rect(fill = "#0d1117", colour = NA),
    panel.grid.major = element_blank(),
    axis.text        = element_text(colour = "grey80"),
    axis.title       = element_text(colour = "grey90"),
    legend.background = element_rect(fill = "#0d1117"),
    legend.text      = element_text(colour = "grey80"),
    legend.title     = element_text(colour = "grey90"),
    plot.title       = element_text(colour = "white", size = 18, face = "bold")
  ) +
  labs(
    title    = "Renewable Energy Share (%) across BRICS",
    subtitle = "Hydro + Wind + Solar contribution to primary energy",
    caption  = "Source: OWID Energy Dataset"
  )



```


